Notes on Prediction Types

### 2011-12-20 docs/pert_notes/pert.tex
update the description of species_pred_type = predict_rhoprime_and_rhoX to be consistent with the recent change.

A piecewise linear prediction of $\phi$ to the interface 
would appear as:
\begin{eqnarray}
\phi_{i+1/2,j}^{n+1/2} &=& \phi_{i,j}^n 
    + \left . \frac{\Delta x}{2} \frac{\partial \phi}{\partial x} \right |_{i,j}
    + \left . \frac{\Delta t}{2} \frac{\partial \phi}{\partial t} \right |_{i,j} \\
 &=& \phi_{i,j}^n 
    + \left . \frac{\Delta x}{2} \frac{\partial \phi}{\partial x} \right |_{i,j}
    +  \frac{\Delta t}{2} \left ( -u \frac{\partial \phi}{\partial x} 
                                         -v \frac{\partial \phi}{\partial y} + f \right ) \\
 &=& \phi_{i,j}^n + \frac{\Delta x}{2} \left ( 1 - \frac{\Delta t}{\Delta x} u \right ) 
           \frac{\partial \phi}{\partial x} 
    \underbrace{- \frac{\Delta t}{2} v \frac{\partial \phi}{\partial y}}_{\mathrm{``transverse~term''}} + \frac{\Delta t}{2} f
\end{eqnarray}
(see the Godunov notes section for more details).  Here, the
``transverse term'' is accounted for in {\tt make\_edge\_scal}.  Any
additional forces should be added to $f$.  For the perturbation form
of equations, we add additional advection-like terms to $f$ by calling
{\tt modify\_scal\_force}.  This will be noted below.

@@

A conservative equation takes the form:
\begin{equation}
\frac{\partial \phi}{\partial t} = -\nabla \cdot ( \phi \Ub) + f
\end{equation}
%
Now a piecewise linear prediction of $\phi$ to the interface is 
\begin{eqnarray}
\phi_{i+1/2,j}^{n+1/2} &=& \phi_{i,j}^n 
    + \left . \frac{\Delta x}{2} \frac{\partial \phi}{\partial x} \right |_{i,j}
    + \left . \frac{\Delta t}{2} \frac{\partial \phi}{\partial t} \right |_{i,j} \\
 &=& \phi_{i,j}^n 
    + \left . \frac{\Delta x}{2} \frac{\partial \phi}{\partial x} \right |_{i,j}
    +  \frac{\Delta t}{2} \left ( -\frac{\partial (\phi u)}{\partial x} 
                                  -\frac{\partial (\phi v)}{\partial y} + f \right ) \\
 &=& \phi_{i,j}^n + \frac{\Delta x}{2} \left ( 1 - \frac{\Delta t}{\Delta x} u \right ) 
           \frac{\partial \phi}{\partial x} 
    \underbrace{- \frac{\Delta t}{2} \phi \frac{\partial u}{\partial x} 
                - \frac{\Delta t}{2} \frac{\partial (\phi v)}{\partial y}}_{\mathrm{``transverse~term''}} + \frac{\Delta t}{2} f
\end{eqnarray}
Here the ``transverse term'' includes the non-advective portion of the $x$-flux (for the $x$-prediction).
Both of the underbraced terms are accounted for in {\tt make\_edge\_scal} automatically when
we call it with {\tt is\_conservative = .true.}.

@@

Here we wish to construct $(\rho X_k)_\mathrm{edge}$ by predicting
$(\rho X_k)$ to the edges as a single quantity.  We start with
Eq.~\ref{species equation}:
\begin{equation}
\frac{\partial(\rho X_k)}{\partial t} =
  -\Ub\cdot\nabla(\rho X_k) \, \underbrace{-\rho X_k \nabla\cdot\Ub}_{``\rho X_k ~\text{force''}} + \rho \omegadot_k \, . \label{species equation labeled}
\end{equation}
The edge state is created by calling {\tt make\_edge\_scal} in {\tt
  density\_advance} with the underbraced term in (\ref{species
  equation labeled}) as the forcing.  This advective force is computed
by {\tt modify\_scal\_force} with {\tt fullform = .true.}.  We do not
consider the $\omegadot_k$ term in the forcing when Strang-splitting.

@@

\noindent {\tt species\_pred\_type = predict\_rhoprime\_and\_rhoX}:
\begin{equation}
\frac{\partial\rho X_k}{\partial t} = 
-\nabla\cdot\left\{\left[\left(\rho X_k \right)_\mathrm{edge} \right](\Ubt+w_0\eb_r)\right\}.
\end{equation}

>>>>

A piecewise linear prediction of $\phi$ to the interface 
would appear as:
\begin{eqnarray}
\phi_{i+1/2,j}^{n+1/2} &=& \phi_{i,j}^n 
    + \left . \frac{\Delta x}{2} \frac{\partial \phi}{\partial x} \right |_{i,j}
    + \left . \frac{\Delta t}{2} \frac{\partial \phi}{\partial t} \right |_{i,j} \\
 &=& \phi_{i,j}^n 
    + \left . \frac{\Delta x}{2} \frac{\partial \phi}{\partial x} \right |_{i,j}
    +  \frac{\Delta t}{2} \left ( -u \frac{\partial \phi}{\partial x} 
                                         -v \frac{\partial \phi}{\partial y} + f \right ) \\
 &=& \phi_{i,j}^n + \frac{\Delta x}{2} \left ( 1 - \frac{\Delta t}{\Delta x} u \right ) 
           \frac{\partial \phi}{\partial x} 
    \underbrace{- \frac{\Delta t}{2} v \frac{\partial \phi}{\partial y}}_{\text{``transverse~term''}} + \frac{\Delta t}{2} f
\end{eqnarray}
(see the Godunov notes section for more details).  Here, the
``transverse term'' is accounted for in {\tt make\_edge\_scal}.  Any
additional forces should be added to $f$.  For the perturbation form
of equations, we add additional advection-like terms to $f$ by calling
{\tt modify\_scal\_force}.  This will be noted below.

@@

A conservative equation takes the form:
\begin{equation}
\frac{\partial \phi}{\partial t} = -\nabla \cdot ( \phi \Ub) + f
\end{equation}
%
Now a piecewise linear prediction of $\phi$ to the interface is 
\begin{eqnarray}
\phi_{i+1/2,j}^{n+1/2} &=& \phi_{i,j}^n 
    + \left . \frac{\Delta x}{2} \frac{\partial \phi}{\partial x} \right |_{i,j}
    + \left . \frac{\Delta t}{2} \frac{\partial \phi}{\partial t} \right |_{i,j} \\
 &=& \phi_{i,j}^n 
    + \left . \frac{\Delta x}{2} \frac{\partial \phi}{\partial x} \right |_{i,j}
    +  \frac{\Delta t}{2} \left ( -\frac{\partial (\phi u)}{\partial x} 
                                  -\frac{\partial (\phi v)}{\partial y} + f \right ) \\
 &=& \phi_{i,j}^n + \frac{\Delta x}{2} \left ( 1 - \frac{\Delta t}{\Delta x} u \right ) 
           \frac{\partial \phi}{\partial x} 
    \underbrace{- \frac{\Delta t}{2} \phi \frac{\partial u}{\partial x} }_{\text{``non-advective~term''}}
                \underbrace{- \frac{\Delta t}{2} \frac{\partial (\phi v)}{\partial y}}_{\text{``transverse~term''}} + \frac{\Delta t}{2} f
\end{eqnarray}
Here the ``transverse term'' is now in conservative form, and an additional
term, the non-advective portion of the
$x$-flux (for the $x$-prediction) appears.  Both of the underbraced terms are
accounted for in {\tt make\_edge\_scal} automatically when we call it
with {\tt is\_conservative = .true.}.

@@

Here we wish to construct $(\rho X_k)_\mathrm{edge}$ by predicting
$(\rho X_k)$ to the edges as a single quantity.  We recall
Eq.~\ref{species equation}:
\begin{equation}
\frac{\partial(\rho X_k)}{\partial t} =
  -\nabla \cdot (\rho \Ub X_k) + \rho \omegadot_k \, . \nonumber
\end{equation}
The edge state is created by calling {\tt make\_edge\_scal} in {\tt
  density\_advance} with {\tt is\_conservative = .true.}.
We do not consider the $\rho \omegadot_k$ term in the forcing when
Strang-splitting.

@@

\noindent {\tt species\_pred\_type = predict\_rhoprime\_and\_rhoX}:
\begin{equation}
\frac{\partial\rho X_k}{\partial t} = 
-\nabla\cdot\left\{\left[\left(\rho X_k \right)_\mathrm{edge} \right](\Ubt+w_0\eb_r)\right\}.
\end{equation} \\
